Domain Specific Language for high performance computing

### 2024-02-09 paper/paper.tex
writting language specification

\documentclass[ams, a4j]{U-AizuGT}
\usepackage{pifont}
% \usepackage{graphicx}
\usepackage{cite}
\usepackage{mathtools}
% \usepackage{listings,jvlisting}
% \documentclass[a4j]{jarticle} %ここは関係ない
\usepackage{listings,jvlisting} 
\usepackage[dvipdfmx]{graphicx}
\usepackage{physics}
\bibliographystyle{ieicetr}

@@

宇宙物理学や流体力学等、様々な分野で粒子系シミュレーションが行われている.
例えば、惑星間の重力による力学的進化をシミュレーションしたN体シミュレーションや
水の流体運動のモデル化がある。
これらのシミュレーションには粒子の数に対して計算量が膨大になる。
例としてN体シミュレーションの場合を説明する。N体シミュレーションでは,
ほとんどの計算時間を重力相互作用の計算が占めている。\eqref{eq:gravity} は、
重力相互作用の式である.
\begin{equation}
a_i = \sum_{\substack{j \neq i}}^{N} G m_j \frac{\vb{x}_j - \vb{x}_i}{\left( \left| \vb{x}_j - \vb{x}_i \right| + \epsilon^2 \right)^{3/2}\label{eq:gravity}}
\end{equation}

ここで $a_i$, $\vb{x}_i$はそれぞれ粒子iの加速度と位置を表し、
$m_j$,$\vb{x}_j$はそれぞれ粒子jの質量と位置を表し、
$N$,$G$,$\epsilon$はそれぞれ粒子の数、重力定数発散を防ぐためのソフトニングパラメータである。
\eqref{eq:gravity}から計算量は粒子数Nに比例して、\begin{math}O(N^2) \end{math}となる。
N体シミュレーションでは、粒子の数が膨大になる場合が多い。例えば、
球状星団という天体には星が約$10^4 ~ 10^5 $ 個存在する。このため非常に長い計算時間がかかるため高速化が求めれる。
粒子間相互作用の計算は、粒子同士の計算が独立しているためデータ並列性が高い。そのため、単一の命令に対して複数の
データを処理するSIMDによる並列に適している。
SIMDによる並列計算の例として、小池らが行ったGRAPE-DR\cite{GRAPE-explain-thesis}による重力多体問題シミュレーションがある。\cite{GRAPE-DR-thesis}
しかし、SIMD計算を行うには,コンピュータのアーキテクチャを考慮したコードを記述しなくてはならない。

この研究の目的は、ユーザーが粒子間相互作用の計算を並列化を意識することなく、並列に実行できるように
することである。そのためにPykerというDomain Specific Language\lparen DSL \rparen を開発した。
Pykerでは、ユーザに必要な変数の定義と数式から相互作用の計算をSIMD命令によって並列化したコードを生成する。
Pykerの評価方法としては、C言語のコンパイラが自動で行うSIMD化とPykerによるSIMD化の性能を比較する。
そのために、3つの相互作用の計算を行うコードをそれぞれにPykerから生成されたSIMD命令を適応しない
場合のコードと、SIMD命令を適応したコードの実行時間を計測し比較する。
また、Pykerを作成するにあたって粒子間相互作用の計算するコードを生成するPIKGを参考にした。

\section{Construction}
We use Sympy, a library in Python for implementing  the implementation of pyker.
SymPy supports various operations, and the formulas using these operations are internally treated as 
syntax trees. We leveraged this functionality to read formulas and generate code capable of parallel execution.
For parallelization, we used an instruction set named Advanced Vector Extensions 2 \lparen AVX2\rparen. AVX2 is one of the 
extended instruction sets implemented in Intel's CPUs. The next section will explain how parallelization is carried out using AVX2.

\subsection{AVX2 for parallelization}
% _mm256dの説明はここですべきか？ codegen で説明すべきか
To perform SIMD operations, just like with regular computations, data is first loaded from memory, computed based on that data,
and the results are stored. However, in the case of SIMD, computations are performed using SIMD-specific registers, so the 
methods of loading, computing, and storing differ. When loading, multiple elements that will undergo the same instruction are
stored in a SIMD-specific register. These are then used to perform computations simultaneously, and the results are stored for
each.

For instance, consider a calculation like C = X + Y, where elements of double-precision floating arrays x and y are computed
. When using AVX2 for SIMD, since the registers are 256 bits, four elements from x and y are loaded each. These are then added 
simultaneously, and the results are stored in the array C. When loading into SIMD registers, if it's a 1-dimensional array, 
contiguous memory access is sufficient, but for data structures like Array of Structures \lparen AoS \rparen shown in Figure a., elements need
to be loaded individually.

For example, when considering the 3-dimensional coordinates of particles as double-precision floats 
in pos\lbrack n\rbrack \lbrack 3\rbrack  \lparen where n is the total number of particles, and pos\lbrack i \rbrack \lbrack 0\rbrack , pos\lbrack i\rbrack \lbrack 1\rbrack , 
pos\lbrack i\rbrack \lbrack 2\rbrack  represent the x, y, z coordinates 
of the i-th particle, respectively\rparen, and calculating the difference in x coordinates between particles i and j, it becomes dx = 
pos\lbrack i\rbrack \lbrack 0\rbrack  - pos\lbrack j\rbrack \lbrack 0\rbrack . However, for parallelization with SIMD, desiring dx0 = pos\lbrack i\rbrack \lbrack 0\rbrack  - pos\lbrack j \rbrack \lbrack 0\rbrack , dx1 = pos\lbrack i+1\rbrack \lbrack 0\rbrack  - pos\lbrack j\rbrack \lbrack 0\rbrack 
, dx2 = pos\lbrack i+2\rbrack \lbrack 0\rbrack  - pos\lbrack j\rbrack \lbrack 0\rbrack , dx3 = pos\lbrack i+3\rbrack \lbrack 0\rbrack  - pos\lbrack j\rbrack \lbrack 0\rbrack , 
the data are packed as \lparen pos\lbrack i\rbrack \lbrack 0\rbrack , pos\lbrack i+1\rbrack \lbrack 0\rbrack , pos\lbrack i+2\rbrack \lbrack  0\rbrack , pos\lbrack  i+3\rbrack \lbrack  0\rbrack \rparen,
\lparen pos\lbrack  j\rbrack \lbrack  0\rbrack , pos\lbrack  j\rbrack \lbrack  0\rbrack , pos\lbrack  j\rbrack \lbrack  0\rbrack , pos\lbrack  j\rbrack \lbrack  0\rparen, \lparen dx0, dx1, dx2, dx3\rparen and subtraction is performed simultaneously. However, since data needs
to be loaded onto SIMD registers all at once, the gather instruction is used to load data for particle i. The gather instruction is used
to read non-contiguous data elements from memory addresses by specifying addresses. An explanation of the gather instruction is shown in 
Figure x. Data loaded in this way are computed simultaneously using SIMD instructions.

\begin{lstlisting}[frame=single]
  dx0 = pos[i][0] - pos[j][0]
  dx1 = pos[i + 1][0] - pos[j][0]
  dx2 = pos[i + 2][0] - pos[j][0]
  dx3 = pos[i + 3][0] - pos[j][0]
\end{lstlisting}

\begin{lstlisting}[frame=single]
  (x1, x2, x3, x4) + (y1, y2, y3, y)

When storing computation results, addresses may 
not be contiguous. In such cases, there is an instruction called scatter, which is the counterpart to gather, but it is not supported 
in AVX2, so data are stored using pointers individually. In this way, parallelization is achieved with SIMD. Moreover, these instructions
can be explicitly handled in high-level programming languages like C and C++ using intrinsic functions. To handle intrinsic functions 
in C and C++, include the file immintrin.h, which is available as a standard in the language. An example of C++ code that calculates the difference in x coordinates, simd\_tmp.cpp, is shown. { Pyker}  generates C++ code capable of executing SIMD instructions like simd\_tmp.cpp
from the described formulas. The next section will explain the Pyker.

@@

>>>>

\documentclass[ams, a4j]{U-AizuGT}
\usepackage{pifont}
% \usepackage{graphicx}
\usepackage{cite}
\usepackage{mathtools}
% \usepackage{listings,jvlisting}
% \documentclass[a4j]{jarticle} %ここは関係ない
\usepackage{listings,jvlisting} 
\usepackage[dvipdfmx]{graphicx}
\usepackage{physics}

\usepackage[utf8]{inputenc}
\usepackage{booktabs} % きれいな表を作成するためのパッケージ

\bibliographystyle{ieicetr}

@@

Particle simulations have been executed in various fields such as 
astrophysics and fluid dynamics. For example, N-body simulations 
simulate the dynamical evolution due to gravity between planets,
and models of fluid motion models simulate water.
Particle simulations are computationally demanding
with respect to the number of particles.
As an example, we will discuss the case of an N-body simulation.
In this simulation, most of the computational time is occupied by
the calculation of the gravitational interaction. equation \eqref{eq:gravity} is 
the equation for the gravitational interaction.

\begin{equation}
a_i = \sum_{\substack{j \neq i}}^{N} G m_j \frac{\vb{x}_j - \vb{x}_i}{\left( \left| \vb{x}_j - \vb{x}_i \right| + \epsilon^2 \right)^{3/2}\label{eq:gravity}}
\end{equation}

$a_i$, $\vb{x}_i$ $m_j$,$\vb{x}_j$,$N$,$G$,$\epsilon$
are the acceleration and position of particle i, the mass of particle j, the number of position particles, respectively, and
softening parameters to prevent gravitational constant divergence.
From equation \eqref{eq:gravity},time complexity is proportional to 
the number of particles N,

 N-body simulations often require large numbers of particles.
For example, a globular cluster contains about $10^4 ~ 10^5 $ stars.
Therefore, very long computation time is required.  parallel computing is 
generally used to speed up the process.

The computation of interparticle interactions has high
data parallelism because the computation of particles is 
independent of each other. Therefore, 
Parallelism by SIMD\lparen Single Instruction/Multiple Data \rparen, 
one of Flynn's taxonomy \cite{flynn1972some} , is suitable.
SIMD means processing multiple data for a single instruction.
However, to perform SIMD, code must be written for the computer architecture.

The goal of this study is to allow users to run particle interaction 
calculations in parallel without having to be aware of parallelization.
We developed a Domain Specific Language\lparen DSL \rparen called Pyker.
DSL is a programming language for solving specific problems.
Pyker generates parallelized code with SIMD instructions for the 
calculation of interactions from the definition of variables and 
the description of formulas.
The evaluation method for Pyker is to compare the performance of 
the automatic SIMD  performed by the C compiler with that of the 
SIMD by Pyker. 
For this purpose, we measured the execution time of the code that
does not adapt the SIMD instructions generated by Pyker for each 
of the three interaction calculations.
The execution time of the code that does not adapt SIMD instructions
generated from Pyker and the code that does adapt SIMD instructions 
are measured and compared.

In this paper, section 2 explains how Pyker defines the interparticle 
interaction equation, section 3 describes the process of code generation 
by Pyker, and section 4 presents the results of the performance evaluation
experiments.

\section{Definition of the interparticle interaction in Pyker}

\subsection{interparticle interaction}
we explain the needed for information to the calculate interparticle 
interaction with reference to the actual code.
gravity.cpp is the C++ source code that performs the calculation 
in equation \eqref{eq:gravity}.
\begin{lstlisting}[frame=single, caption=skeleton code, label=fuga]

  for (int i = 0; i < n; i++) {
    double ax = 0.0;
    double ay = 0.0;
    double az = 0.0;

    for (int j = 0; j < n; j++) {
        double dx = double(xj[j][0] - xi[i][0]);
        double dy = double(xj[j][1] - xi[i][1]);
        double dz = double(xj[j][2] - xi[i][2]);
        double r2 = dx * dx + dy * dy + dz * dz + eps2;
        double ri2 = 1.0f / sqrt(r2);

        double mr = mj[j] * ri2 * ri2 * ri2;

        ax += double(mr * dx);
        ay += double(mr * dy);
        az += double(mr * dz);
    }

    acci[i][0] = ax;
    acci[i][1] = ay;
    acci[i][2] = az;
}
\end{lstlisting}

We will explain how this code corresponds to the expression 
\eqref{eq:gravity}.
acci,xi,mj,xj,eps2 correspond to
 $a_i$, $\vb{x}_i$ $m_j$,$\vb{x}_j$,$N$,$\epsilon$ respectively.
The $G$ is not in the code because it can be calculated after the 
calculation of the interparticle interactions.

 From this code, we can see that the variables to store the results, 
the variables for particle i and particle j respectively, and the 
variables for $N$ and $G$ can be calculated after the calculation of 
the particle-particle interaction. The other variables have different 
processing and data structures.

 Therefore, there are three categories of information in a variable.
\begin{itemize}
  \item Whether the variable is a particle i, a particle j, a
  variable that stores the result, or any other variable
  \item The type information of the variable.
  \item Whether the variable is a vector or not
\end{itemize}

 With the definition of this variable and a description of the
mathematical equation using the variable, a code to calculate the
interparticle interaction can be generated.
これらをもとにPykerの言語仕様を説明する。

\subsection{language specification}
Pykerでは、大きく、変数定義と相互作用記述に分かれる。
まず変数定義について説明する。

 先ほどのセクションよりPykerの変数定義を以下に正規表現で示す。
% \begin{lstlisting}  
\["(EPI|EPJ|FORCE)?\:\text{ } (vec3<)?(F64)>?\:\text{ } \text{\textbackslash w+}"\]
% \end{lstlisting}
EPI,EPJ,FORCEはそれぞれ粒子i,粒子j,結果を保存する変数であることを示す。それ以外の変数は何も書かない
ベクトルはvec3で表す。
もしベクトルであれば'<>'の中に型を定義する。
型は倍精度小数だけでF64で示す。
最後に変数名を記述する。

次に相互作用記述の説明をする。
相互作用の記述に使える演算は以下の図に示す。

以下に\label{eq:gravity}の計算をPyker記述した例を示す。
\begin{lstlisting}[frame=single, caption=gravity\_interparticle, label=fuga]
  EPI vec3<F64> ri
  EPJ vec3<F64> rj
  EPJ F64 mass
  EPJ F64 eps2
  FORCE vec3<F64> ai
  rij = ri - rj
  r2 = rij * rij + eps2
  r_inv = 1.0 / sqrt(r2)
  r2_inv = r_inv * r_inv
  mr_inv = mass * r_inv
  mr3_inv = r2_inv * mr_inv
  ai = mr3_inv * rij
\end{lstlisting}

5行目までは変数定義部になる。

\begin{table}[ht]
  \centering % 表を中央揃えにする
  \caption{サンプル表} % 表のタイトル
  \label{tab:sampleTable}
  \begin{tabular}{ccc} % ここで列の数とテキストの配置を指定
  \toprule
  項目 & 値 & 単位 \\
  \midrule
  項目1 & 100 & kg \\
  項目2 & 200 & g \\
  項目3 & 300 & mg \\
\bottomrule
\end{tabular}
\end{table}

\section{Construction}
Pykerでは、どういうことができるのか？

We use Sympy, a library in Python for implementing  the implementation of pyker.
SymPy supports various operations, and the formulas using these operations are internally treated as 
syntax trees. We leveraged this functionality to read formulas and generate code capable of parallel execution.
For parallelization, we used an instruction set named Advanced Vector Extensions 2 \lparen AVX2\rparen. AVX2 is one of the 
extended instruction sets implemented in Intel's CPUs. The next section will explain how parallelization is carried out using AVX2.

@@

\begin{table}[ht]
  \centering % 表を中央揃えにする
  \caption{サンプル表} % 表のタイトル
  \label{tab:sampleTable}
  \begin{tabular}{ccc} % ここで列の数とテキストの配置を指定
  \toprule
  項目 & 値 & 単位 \\
  \midrule
  項目1 & 100 & kg \\
  項目2 & 200 & g \\
  項目3 & 300 & mg \\
\bottomrule
\end{tabular}
\end{table}
